---
output:
  pdf_document: default
  html_document: default
---
#Assignment 0 SRT411 Github: odamicostudent
The purpose of this assignment was to go through the
todo sections in a short introduction to R at 
https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf


#Section 2.4 Working Directory 
Setting a working directory in R can be done using:
```{r, eval=FALSE}
setwd("~")
```
where "~" can be replaced by the desired directory, you can also use tools --> Set working directory in RStudio

#Section 2.5 Libraries
R, like any language, has many packages. These can be listed using:
```{r}
library()
```
To install and use new packages, use:
```{r, eval=FALSE}
install.packages("geometry")
library("geometry")
```
#Section 3.1 Calculator
Equations can simply be typed to give answers as so:
```{r}
10^2 + 36
```
R will return your answer as seen above.

#Todo task number 1
Percentage of my life spent at college: 
```{r}
(2017 - 2014) / (2014 - 1999) * 100
```

The equals operator can be used to assign values to variables, simply entering the variable name will
tell R to print the value stored in the variable, you can use the variable name in equations and re-assign the values as needed:
```{r}
a = 5
a
(a * 3)^2
a = a *20
a
```
To remove all variables from R's memory, use rm:
```{r, eval=FALSE}
rm(list=ls())
```
To remove specific variables, use their name(s) instead:
```{r, eval=FALSE}
rm(a)
```
#Todo task number 2
Repeating the percentage question using variables:
```{r}
start = 2017
compare_date = 2014
born= 1999
(start - compare_date) / (compare_date - born) * 100
```

#3.3 Scalars, vectors, and functions
Scalars are single values, vectors are a row of numbers like an array.
Before, we created a scalar value, to create a vector, 
use the c(n1, n2, n3, ...n) function:
```{r}
a = c(1, 2, 3, 5, 1234, 3.1)
a
```
to use the 'mean' function we use mean(n) where n is the vector to use.
```{r}
mean(a)
```

#Todo task number 3
```{r}
sum(4, 5, 8, 11)
```

#Todo task number 4
R lets you create random normal distributions using rnorm()
```{r}
rnorm(100)
```

#Todo task number 5
R allows you to access useful help pages using help()
```{r, eval=FALSE}
help(sqrt)
```
This example will output an HTML page with useful information about
the sqrt() function

#Todo task number 6
loading firstscript.R into knitr
```{r cache=FALSE}
knitr::read_chunk('firstscript.R')
```
running the code found at firstscript.R
```{r test-a}
```


#Todo task number 7
You can create a matrix using the syntax matrix(data=v, ncol=n) where v is the 
vector and n is the number of columns to use
```{r}
P = seq(from=31,to=60,by=1)
M = matrix(data=P, ncol=5, dimnames = list(c(1, 2, 3, 4, 5, 6),c("Q", "Q", "Q", "Q", "Q")))
M
```

#Todo task number 8
```{r}
for(j in c(1, 2, 3, 4, 5)){
x1 = (as.integer(runif(100, 0, 1000)))
x2 = (as.integer(runif(100, 0, 1000)))
x3 = (as.integer(runif(100, 0, 1000)))
x2 = x1+x2
x3 = x1+x2+x3
t = data.frame(a= x1,b=x2, c=x3)
plot(t)
rm(list=ls())
}
```

The sd output can be generated using sapply(), sd has been defunct in R for
quite some time and no longer works
```{r}
x1 = (as.integer(runif(100, 0, 1000)))
x2 = (as.integer(runif(100, 0, 1000)))
x3 = (as.integer(runif(100, 0, 1000)))
x2 = x1+x2
x3 = x1+x2+x3
t = data.frame(x1,x2,x3)
sapply(t, sd)
rm(list=ls())
```
running this script a few more times produces the following outputs
```{r, echo=FALSE}
x1 = (as.integer(runif(100, 0, 1000)))
x2 = (as.integer(runif(100, 0, 1000)))
x3 = (as.integer(runif(100, 0, 1000)))
x2 = x1+x2
x3 = x1+x2+x3
t = data.frame(x1,x2,x3)
sapply(t, sd)
rm(list=ls())
```

```{r, echo=FALSE}
x1 = (as.integer(runif(100, 0, 1000)))
x2 = (as.integer(runif(100, 0, 1000)))
x3 = (as.integer(runif(100, 0, 1000)))
x2 = x1+x2
x3 = x1+x2+x3
t = data.frame(x1,x2,x3)
sapply(t, sd)
rm(list=ls())
```

```{r, echo=FALSE}
x1 = (as.integer(runif(100, 0, 1000)))
x2 = (as.integer(runif(100, 0, 1000)))
x3 = (as.integer(runif(100, 0, 1000)))
x2 = x1+x2
x3 = x1+x2+x3
t = data.frame(x1,x2,x3)
sapply(t, sd)
rm(list=ls())
```

```{r, echo=FALSE}
x1 = (as.integer(runif(100, 0, 1000)))
x2 = (as.integer(runif(100, 0, 1000)))
x3 = (as.integer(runif(100, 0, 1000)))
x2 = x1+x2
x3 = x1+x2+x3
t = data.frame(x1,x2,x3)
sapply(t, sd)
rm(list=ls())
```

R also has lists whose elements don't have to be of equal sizes
```{r}
x1 = (as.integer(runif(100, 0, 1000)))
x2 = (as.integer(runif(100, 0, 1000)))
x3 = (as.integer(runif(100, 0, 1000)))
x2 = x1+x2
x3 = x1+x2+x3
t = data.frame(a=x1,b=x2,c=x3)
plot(t$a, type="l", ylim=range(t),
lwd=3, col=rgb(1,0,0,0.3))
lines(t$b, type="s", lwd=2,
col=rgb(0.3,0.4,0.3,0.9)) 
points(t$c, pch=20, cex=4,
col=rgb(0,0,1,0.3))
```
rgb is "RedGreenBlue", it allows us to choose colors for 
our points. lwd is line width. pch is the type of symbol to use,
R has various symbols that can be used for points which are indexed
by numbers, 20 is a dot symbol. cex is the scale of the symbols

#todo task number 9
```{r}
d = read.table(file="tst2.txt", header = TRUE)
d
d$g = d$g*5
d
```

#todo task number 10
```{r}
mean(sqrt(rnorm(100)))
```
I get an error, therefore, NA has to be inserted where data is
unavailable. If you want to ignore NAs just use na.rm=TRUE as an 
argument to ignore them.

#todo task number 11
```{r}
dates=strptime(c(20190215, 20141205, 20190903), format="%Y%m%d") 
presents=c(0,0,5)
df <- data.frame(dates, presents)
plot(df)
```


#todo task number 12
```{r}
n = seq(from=1, to=100, by=1)
f = c()
for(x in n){
if(x < 5 | x > 90){
f[x] = n[x] * 10
}
else{
f[x] = n[x] * 0.1
}  
  
}
f
```

#todo task number 13
```{r}
func = function(arg1){
f=c()  
for(i in arg1){
if(i < 5 | i > 90){
f[i] = arg1[i] * 10
}
else{
f[i] = arg1[i] * 0.1
}  
}
f
}
func(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10))

```

#todo task number 14
Here, I have to prove the following footnote
"... people  often  use  more  for-loops  than  nec-essary.The  ToDo  above  can  be  done  more  easilyand quickly without a for-loop but with regular vector-computations."
```{r}
x=c(1, 2, 30, 4, 10, 20, 90, 123, 14567)
x
f <- ifelse(x<5,x*10, ifelse(x>90, x*10, x*0.1))
f
```
No For loop was needed here













